---
title: "exercise-04"
output: html_document
date: "2024-02-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Instructions: <https://difiore.github.io/ada-2024/exercise-04.html>

```{r}
library(curl)
f <- curl("https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/data-wrangling.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

## Wordle Rules

The rules of **Wordle** are simple: A player has SIX tries to guess a 5-letter word that has been selected at random from a list of possible words. Guesses need to be valid 5-letter words. After each guess, a player receives feedback about how close their guess was to the word, which provides information they can use to try to solve the puzzle. This feedback lets you know whether each letter your guess is either [1] *in the solution word and in the correct spot*, [2] *in the solution word but in the wrong position*, or [3] *not in the solution word*. In app/browser versions of **Wordle**, this feedback is provided visually using colors, but that need not be the case.

Preliminaries

-   Set up a new ***GitHub*** repo in you or your partner\'s ***GitHub*** workspace named \"exercise-04\" and clone that down to both of your computers as a new ***RStudio*** project. The instructions outlined as **Method 1** in [**Module 6**](https://difiore.github.io/ada-2024/06-module.html) will be helpful. Make sure that both you and your partner are collaborators on the repo and that you add me as a collaborator as well (my ***GitHub*** username is \"difiore\").

-   Go to <https://github.com/difiore/ada-2024-datasets>, download the following two data files, and add them to your repo:

    -   collins-scrabble-words-2019.txt

    -   google-10000-english-usa-no-swears.txt

```{r download} # not sure if ths actually needs a block...}
library(tidyverse)
read_tsv("https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/google-10000-english-usa-no-swears.txt", col_names = TRUE)
read_tsv("https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/collins-scrabble-words-2019.txt", col_names = TRUE)


read_csv("collins-scrabble-words-2019.txt")
```

The first of these files (279,497 lines long) contains a list of \"Official Scrabble Words\" in the English language based on the ***Collins English Dictionary*** published by HarperCollins. The first line in the file is the word \"words\" and is used as a header.

The second of these files (9885 lines long) contains a list of \~10,000 of the most common words in the English language, based on data compiled by Google, and omitting common swear words. Again, the first line is the word \"words\" and is used as a header.

#### **Step 1**

-   DONE: Create your own custom function called `load_dictionary()` that takes a single argument, \"filename\", that can be used to read in either of the two data files your downloaded.

```{r load_dictionary fxn}
load_dictionary <- function(filename) {
  as.vector(read_csv(filename))
  # convert "words" to list
}
```

-   DONE: Once you have created your function, use that function to create two variables, **solution_list** and **valid_list**, that, respectively contain *vectors* of possible solution words and valid words to guess.

```{r lists}
solution_list <- load_dictionary("collins-scrabble-words-2019.txt")
valid_list <- load_dictionary("google-10000-english-usa-no-swears.txt")

# Running str(valid_list) should return… [chr [1:279496] "AA" "AAH" "AAHED" "AAHING"...]
str(valid_list) # looks good

# Running str(solution_list) should return…[chr [1:8336] "THE" "OF" "AND" "TO"...]
str(solution_list) # looks good
```

#### **Step 2**

DONE: Winnow your variable **solution_list** to only include words that are included in **valid_list**, using `intersect()`. How many words are in your updated **solution_list** vector?

```{r winnow_lists}
#doesn't work if I load in as vector...
valid_sol_list <- intersect(solution_list, valid_list)
# 8336 words
```

#### **Step 3**

Write a custom function called `pick_solution()` that [1] removes all words from **solution_list** that are not 5 letters in length, [2] then randomly chooses a single word from those that remain, and [3] then splits that word into a vector of single-character elements. You should be able to pass your **solution_list** vector as the argument to the function.

```{r pick_sol_fxn}

#pass in data frame
pick_solution <- function() {
  # 1. remove non-5 letter words
  five_letter_words <- solution_list |>
    mutate(letters = ... )
    filter(solution_list,...) # where letters = 5) |>
    mutate(letters = NULL) |>
    # convert to list

  # randomly chooses word from 5-letter word list
    correct_word <- sample(five_letter_words)
  
  #3 split word into a vector
  characters <- as.vector(strsplit(correct_word, ""))
}
```

> **HINT**: For [1], you will want to subset the `solution_list` vector according to some criterion of word length. For [2], you may find the `sample()` function useful (use ***R*** help to look up documentation on that function). For [3], you may want to look at the functions `strsplit()` from {base} ***R*** or `str_split()` from the {stringr} package (part of {tidyverse}). Pay attention to the data structure that those functions return, because you will need to carefully reference one of the elements that is returned in order to wind up with a vector.

As a bonus, you might include a second argument for your `pick_solution()` function called \"word_length\" that makes your function flexible enough to select a solution word that is something other than 5 characters long.

Once your function works, run it and assign the result to a variable called **solution**.

```         
solution <- pick_solution(solution_list)
```

#### **Step 4**

Now, to tackle the bulk of the problem, create two more functions. The first should be called `play_wordle()` and it should take three arguments: [1] the answer to the puzzle (the value of your **solution** variable), [2] a list of valid guesses (the contents of your **valid_list** variable), and [3] a value for \"number of guesses\", which you should set to the original **Wordle** game default of 6.

> **HINT**: Here is possible skeleton code for that function.

```         
play_wordle <- function(solution, valid_list, num_guesses=6){
      <function code here>
    }
```

Think carefully about what the `play_wordle()` function needs to do. It should:

1.  At the onset, tell the player the rules of the game, e.g., \"You have ... chances to guess a word of length ...\"

2.  Display what letters the player has not yet guessed (at the onset, this would be all 26 letters of the alphabet), e.g., \"Letters left: ...\"

> **HINT:** There is a built-in dataset in ***R*** called `LETTERS`, which contains the 26 capital letters in the English alphabet and another, `letters`, that contains all the lowercase letters. You will probably want to use either the `toupper()` or `tolower()` functions to ensure that you are always working with the same formatted letters in words and guesses.

```         
# using the LETTERS built-in vector
LETTERS
```

```         
##  [1] "A" "B" "C" "D" "E" "F" "G" "H" "I" "J" "K" "L" "M" "N" "O" "P" "Q" "R" "S"
## [20] "T" "U" "V" "W" "X" "Y" "Z"
```

```         
# using the `toupper()` function
toupper("change my case")
```

```         
## [1] "CHANGE MY CASE"
```

3.  Prompt the player for a guess, e.g., \"Enter guess number ...\", read in their guess, and check that their guess is valid (i.e., that it contains the correct number of letters and is a word included in the \"valid\" word list).

> **HINT:** The `readline()` function, which can take a character string as an argument, will provide a \"prompt\" entering a line of numeric or character data. Hitting `<enter>` or `<return>` signals the end of the line.

```         
guess <- readline("Enter some data here, then press <enter>: ")
```

4.  Compare the guess to the solution word and generate the necessary feedback, e.g., `*` for *in the word and in the correct position*, `+` for *in the word but in the wrong position*, and `-` for *not in the word*. For this step, try writing a separate \"helper\" function, `evaluate_guess()`, called from within `play_wordle()`. This function should take, as arguments, the player\'s guess and the value of the **solution** variable. This is probably the trickiest part of the problem to program, and there are lots of approaches you might take to evaluating guesses. After you work on this for a while, I can share one solution.

5.  Update the list of letters not yet guessed.

> **HINT**: Again, consider using set operations to update the list of letters not yet guessed. `setdiff()` is a function that returns the difference between two vectors.

6.  Check if the puzzle was solved. If so, the function should indicate that the player WON the game and print out their guess and feedback history. If not, the function should prompt the player for another guess, unless they have already hit the maximum number of guesses allowed.

7.  If all guesses are exhausted, the function should indicate that the player LOST the game and, again, print out their guess and feedback history.
