---
title: "exercise-04"
output: html_document
date: "2024-02-12"
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Instructions: <https://difiore.github.io/ada-2024/exercise-04.html>

# Wordle Program

## Preliminaries

-   Set up a new ***GitHub*** repo - "exercise-04"

-   Download data files, and add them to your repo:

    -   collins-scrabble-words-2019.txt (list of official scrabble words)

    -   google-10000-english-usa-no-swears.txt (\~10,000 most common english words)

```{r download}
library(tidyverse)

scrabble_wrds <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/google-10000-english-usa-no-swears.txt"
google_words <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/collins-scrabble-words-2019.txt"
```

## **Step 1:** Create load_dictionary() function & word lists

```{r load_dictionary fxn}
load_dictionary <- function(filename) {
  as.character(unlist(read_tsv(filename, col_names = TRUE)))}

# create solution_list (possible solution words) and valid_list (valid words to guess)
solution_list <- load_dictionary(scrabble_wrds)
valid_list <- load_dictionary(google_words)

# check class
str(valid_list) # looks good
str(solution_list) # looks good

rm(scrabble_wrds, google_words) # clear from environment
```

## **Step 2: Winnow solution_list to only include valid words**

```{r winnow_lists}
solution_list <- intersect(solution_list, valid_list) # 8336 words
```

## **Step 3: Write pick_solution() fxn**

```{r pick_sol_fxn, warning = FALSE}
pick_solution <- function(solution_list, word_length = 5) {
  five_letter_words <- subset(solution_list, nchar(solution_list) == word_length) # create subset of 5-letter words
  solution <- sample(five_letter_words, 1, replace = FALSE) # choose random word from subset
  solution_char <- str_split_1(solution, "") # splits random word into vector of characters   
  return(solution_char)
}

solution <- pick_solution(solution_list)
```

## **Step 4: Create play_wordle() and evaluate_guess() fxns**

1.  `play_wordle()` : pass in (1) answer to the puzzle (**solution** variable), (2) list of valid guesses (**valid_list** variable), (3) value for "number of guesses", defaulting to 6.
    -   At the onset, print rules of the game.
    -   Display unguessed letters.
    -   Prompt the player for a guess, read in their guess, and check that their guess is valid using evaluate_guess() sub-fxn (to check if guess has 5 letters and is included in the "valid" word list).
    -   Compare the guess to the solution word and generate feedback
    -   Update & print the list of letters not yet guessed
    -   Check if the puzzle was solved, and display whether the player WON and print out their guess and feedback history. If not, prompt the player for another guess, unless they've already hit 6 guesses.
    -   After 6 guesses, indicate that the player LOST the game and, again, print out their guess and feedback history.

```{r play_wordle}
play_wordle <- function(solution, valid_list, num_guesses = 6) {
  
  # evaluate guess sub-fxn 
  evaluate_guess <- function(valid_list) {
    guess <- "" # initialize empty val
    guess_validate <- FALSE # initialize boolean as FALSE to enter loop
    
    while (!guess_validate) {
      guess <- toupper(readline("Enter a 5-letter word, then press <enter>: "))
      
      if (nchar(guess) == 5 && guess %in% valid_list) {
        guess_validate <- TRUE  # change validation boolean, exit loop
      } else {
        invalid_msg <- "Invalid guess. Try again."
        print(invalid_msg)
      }
    }
    return(guess)
  }
  
  # print instructions 
  print(c("You have six tries to guess a random 5-letter word.",
          "Guesses need to be valid 5-letter words.",
          "After each guess, you will be told whether each letter your guess is either",
          "[1] in the solution word and in the correct spot,",
          "[2] in the solution word but in the wrong position, or",
          "[3] not in the solution word."))
  
  # create output table
  result_output <- matrix(nrow = 0, ncol = 5, dimnames = list(c(), c("L1", "L2", "L3", "L4", "L5")))
  
  # prompt guess & read in
  guessed_letters <- list() # initialize guessed letters list
  
  for (i in 1:num_guesses) { 
    # print unguessed letters
    unguessed_letters <- setdiff(LETTERS, guessed_letters)
    print(paste("Letters left:", paste(unguessed_letters, collapse = " ")))
    
    # read in and evaluate user guess
    guess <- evaluate_guess(valid_list) # send to validation sub-fxn
    guess_char <- str_split(guess, "")[[1]] # split guess into vector of characters
    guessed_letters <- unique(c(guessed_letters, guess_char)) # add letters to list of guessed letters
    
    result <- guess_char
    for (x in seq_along(guess_char)) {
      if (guess_char[x] %in% solution) {
        result[x] <- paste0(guess_char[x], "*")
      }
      if (guess_char[x] == solution[x]) {
        result[x] <- paste0(guess_char[x], "^")
      }
    }
    
    # add to output table
    result_output <- rbind(result_output, result)
    
    # print outputs
    print(paste("Guess", i, "of", num_guesses))
    print(result_output)
    print(c("Result indicators:",
            "^ : in solution and in correct position",
            "* : in solution but in wrong position", 
            "- : not in the solution word"))
    
    if (all(guess_char == solution)) {
      print("Congratulations! You won!")
      return(invisible())
    }
  }
  print("Try again!")
  
  restart <- toupper(readline("New Game? (Y/N): "))
  if (restart[1] == "Y") {
    play_wordle()
  } else {
    print("Thanks for playing!")
  }
}
```
